Manuscript 1 code

### 2024-11-27 manuscript_1_code.Rmd
Updated stat info for new species changes discovered in Nov 2024 by J Renzi

cover_means <- cover_df %>%
  group_by(year, habitat) %>%
  summarise(macroalgae = mean(macroalgae),
            coral = mean(coral),
            turf = mean(turf),
            cca = mean(cca),
            sand = mean(sand),
            other = mean(other)
            )

cover_se <- cover_df %>%
  group_by(year, habitat) %>%
  summarise(macroalgae = se(macroalgae),
            coral = se(coral),
            turf = se(turf),
            cca = se(cca),
            sand = se(sand),
            other = se(other)
            )
```

@@

taxa_sum_check <- macro_functional_groups %>%
  group_by(location, year) %>%
  summarise(sum_macrocover = sum(prop_cover))

@@

#summarizing functional groups at the quad level
fg_summary <- macro_functional_groups %>%
  group_by(location, year, functional_group, habitat, site) %>%
  summarise(sum_fg_cover = sum(prop_cover))

@@

#summarizing functional groups at the site level
#NOTE that this is ADDED and not averaged. This data is only used to find richness at the site level. LE did this as adding and as a sum, both end up the same from a richness perspective. 
fg_summary_site_SUM <- macro_functional_groups %>%
  group_by(site, year, habitat, functional_group) %>%
  summarise(sum_fg_cover = sum(prop_cover))

#QAQC
fg_summary_qaqc <- macro_functional_groups %>%
  group_by(location, year, functional_group) %>%
  summarise(sum_fg_cover = sum(prop_cover))

fg_qaqc <- fg_summary_qaqc %>%
  group_by(location, year) %>%
  summarise(sum_fg_check = sum(sum_fg_cover))

@@

#pivot wide at quad level
fg_summary_wide <- fg_summary %>%
  pivot_wider(names_from = "functional_group", values_from = "sum_fg_cover")

@@

cover_mod <- glmmTMB(cover_trans ~ richness*habitat + (1|site/location) + (1|year), family = beta_family(), data = alpha_diversity_quad_macro)
summary(cover_mod)
Anova(cover_mod)
#                     Chisq Df Pr(>Chisq)    
# richness         30916.73  1  < 2.2e-16 ***
# habitat            141.78  3  < 2.2e-16 ***
# richness:habitat   294.54  3  < 2.2e-16 ***
hist(residuals(cover_mod)) # looks fine
plot(residuals(cover_mod) ~ fitted(cover_mod)) # negative trend but it can't be less than 0 so not too concerned. some wave pattern
r.squaredGLMM(cover_mod) # 0.631, 0.650
performance::r2(cover_mod) # 0.634, 0.654
emtrends(cover_mod, pairwise ~ habitat, var = "richness") # backreef not different from forereef, forereef also not different
# contrast                    estimate     SE  df z.ratio p.value
# Fringing - Backreef          0.34720 0.0241 Inf  14.409  <.0001
# Fringing - Forereef 10m      0.34865 0.0240 Inf  14.514  <.0001
# Fringing - Forereef 17m      0.38953 0.0250 Inf  15.599  <.0001
# Backreef - Forereef 10m      0.00144 0.0205 Inf   0.071  0.9999
# Backreef - Forereef 17m      0.04232 0.0214 Inf   1.979  0.1957
# Forereef 10m - Forereef 17m  0.04088 0.0209 Inf   1.952  0.2063
```

@@

summary(cover_mod_fg)
car::Anova(cover_mod_fg)
#                              Chisq Df Pr(>Chisq)    
# fg_richness_quad         31341.052  1  < 2.2e-16 ***
# habitat                     49.887  3  8.444e-11 ***
# fg_richness_quad:habitat   447.780  3  < 2.2e-16 ***
hist(residuals(cover_mod_fg)) # good

r.squaredGLMM(cover_mod_fg) # 0.64, 0.65
performance::r2(cover_mod_fg) # 0.63, 0.65
emtrends(cover_mod_fg, pairwise ~ habitat, var = "functional_richness") 
# contrast                    estimate     SE  df z.ratio p.value
# Fringing - Backreef            0.214 0.0266 Inf   8.051  <.0001
# Fringing - Forereef 10m        0.346 0.0261 Inf  13.252  <.0001
# Fringing - Forereef 17m        0.530 0.0262 Inf  20.244  <.0001
# Backreef - Forereef 10m        0.132 0.0231 Inf   5.708  <.0001
# Backreef - Forereef 17m        0.316 0.0232 Inf  13.617  <.0001
# Forereef 10m - Forereef 17m    0.184 0.0223 Inf   8.241  <.0001

@@

```{r synchrony cover mod}
synch_cov_mod <- glmmTMB(cover_mean ~ synchrony*habitat + (1|site), family = beta_family(), data = diversity_stability_synchrony)
summary(synch_cov_mod)
Anova(synch_cov_mod)
#                     Chisq Df Pr(>Chisq)    
# synchrony         209.545  1  < 2.2e-16 ***
# habitat            52.512  3  2.330e-11 ***
# synchrony:habitat  54.632  3  8.227e-12 ***
hist(residuals(synch_cov_mod)) # slight skew
plot(residuals(synch_cov_mod) ~ fitted(synch_cov_mod)) # same issues as always

# can't compute r2 - manually fit null model. Null model won't fit without sv_trans() for some reason
null_synch_mod <- glmmTMB(sv_trans(cover_mean) ~ 1, family = beta_family(), data = diversity_stability_synchrony)
performance::r2(model = synch_cov_mod, null_model = null_synch_mod)
emtrends(synch_cov_mod, pairwise ~ habitat, var = "synchrony") # forereef 17 distinct from all other habitats

@@

### model at site level
```{r site richness mod}
site_cover_mod <- glmmTMB(cover_trans ~ richness*habitat + (1|site_habitat) + (1|year), family = beta_family(), data = alpha_diversity_site_macro)
summary(site_cover_mod)
Anova(site_cover_mod)
#                    Chisq Df Pr(>Chisq)    
# richness         58.9195  1  1.643e-14 ***
# habitat           6.6247  3    0.08487 .  
# richness:habitat  3.2967  3    0.34811    
hist(residuals(site_cover_mod)) # some positive skew, not the worst thing ever
plot(residuals(site_cover_mod) ~ fitted(site_cover_mod)) # a little heteroscedastic but can't be less than 0
r.squaredGLMM(site_cover_mod) # 0.2212534, 0.5990825
performance::r2(site_cover_mod) # 0.22, 0.59
emtrends(site_cover_mod, pairwise ~ habitat, var = "richness") # no difference between the habitats
# contrast                    estimate     SE  df z.ratio p.value
# Fringing - Backreef          -0.0163 0.0394 Inf  -0.413  0.9762
# Fringing - Forereef 10m       0.0639 0.0503 Inf   1.271  0.5818
# Fringing - Forereef 17m       0.0334 0.0612 Inf   0.546  0.9476
# Backreef - Forereef 10m       0.0801 0.0460 Inf   1.742  0.3020
# Backreef - Forereef 17m       0.0497 0.0583 Inf   0.853  0.8289
# Forereef 10m - Forereef 17m  -0.0304 0.0651 Inf  -0.467  0.9662
```

@@

#                            Chisq Df Pr(>Chisq)    
# fg_richness_site         30.9412  1   2.66e-08 ***
# habitat                   4.8929  3     0.1798    
# fg_richness_site:habitat  4.2193  3     0.2387

hist(residuals(site_cover_mod_fg)) # positive skew
plot(residuals(site_cover_mod_fg) ~ fitted(site_cover_mod_fg)) # a little heteroscedastic but can't be less than 0
r.squaredGLMM(site_cover_mod_fg) # 0.180, 0.563
performance::r2(site_cover_mod_fg) # 0.180, 0.563
emtrends(site_cover_mod_fg, pairwise ~ habitat, var = "functional_richness") # no difference between the habitats
# Fringing - Backreef           0.0379 0.0825 Inf   0.459  0.9679
# Fringing - Forereef 10m       0.1846 0.0930 Inf   1.985  0.1935
# Fringing - Forereef 17m       0.0933 0.0960 Inf   0.972  0.7655
# Backreef - Forereef 10m       0.1467 0.0986 Inf   1.488  0.4447
# Backreef - Forereef 17m       0.0554 0.1018 Inf   0.545  0.9480
# Forereef 10m - Forereef 17m  -0.0913 0.1093 Inf  -0.835  0.8377

@@

```{r synchrony cover mod}
synch_cov_mod_site <- glmmTMB(cover_mean ~ synchrony*habitat, family = beta_family(), data = diversity_stability_synchrony_site)
summary(synch_cov_mod_site)
Anova(synch_cov_mod_site)
#                    Chisq Df Pr(>Chisq)  
# synchrony         2.4184  1    0.11992  
# habitat           7.8002  3    0.05033 .
# synchrony:habitat 8.8866  3    0.03084 *
performance::r2(synch_cov_mod_site)
hist(residuals(synch_cov_mod_site)) # slight skew
plot(residuals(synch_cov_mod_site) ~ fitted(synch_cov_mod_site)) # looks fine
emtrends(synch_cov_mod_site, pairwise ~ habitat, var = "synchrony") # forereef 17 different from backreef
# contrast                    estimate   SE  df z.ratio p.value
# Fringing - Backreef           -4.299 3.38 Inf  -1.270  0.5821
# Fringing - Forereef 10m       -0.279 3.22 Inf  -0.087  0.9998
# Fringing - Forereef 17m        3.675 3.78 Inf   0.973  0.7649
# Backreef - Forereef 10m        4.019 1.98 Inf   2.031  0.1764
# Backreef - Forereef 17m        7.973 2.80 Inf   2.848  0.0228
# Forereef 10m - Forereef 17m    3.954 2.59 Inf   1.526  0.4217
```

@@

```{r richness stability mod}
rich_stab_mod <- glmmTMB(cover_stability~ richness_mean*habitat + (1|site), family = Gamma(link = "log"), data = diversity_stability_synchrony)
summary(rich_stab_mod)
Anova(rich_stab_mod)
#                          Chisq Df Pr(>Chisq)    
# richness_mean         2721.438  1  < 2.2e-16 ***
# habitat                325.876  3  < 2.2e-16 ***
# richness_mean:habitat   33.926  3  2.054e-07 ***
hist(residuals(rich_stab_mod)) # looks good
plot(residuals(rich_stab_mod) ~ fitted(rich_stab_mod)) # heteroskedastic - bring to group
r.squaredGLMM(rich_stab_mod) # 0.803, 0.808
performance::r2(rich_stab_mod) # 0.812, 0.816
emtrends(rich_stab_mod, pairwise ~ habitat, var = "richness_mean") # fringing not diff from outer17, backreef not different from outer10
# contrast                     estimate     SE  df z.ratio p.value
# Fringing - Backreef          0.186670 0.0414 Inf   4.512  <.0001
# Fringing - Forereef 10m      0.187196 0.0455 Inf   4.114  0.0002
# Fringing - Forereef 17m      0.027112 0.0463 Inf   0.586  0.9364
# Backreef - Forereef 10m      0.000525 0.0405 Inf   0.013  1.0000
# Backreef - Forereef 17m     -0.159559 0.0411 Inf  -3.885  0.0006
# Forereef 10m - Forereef 17m -0.160084 0.0439 Inf  -3.647  0.0015
```

@@

#                             Chisq Df Pr(>Chisq)    
# fg_richness_mean         3079.229  1  < 2.2e-16 ***
# habitat                   118.605  3  < 2.2e-16 ***
# fg_richness_mean:habitat   25.692  3  1.107e-05 ***

hist(residuals(rich_stab_mod_fg)) # looks fine
plot(residuals(rich_stab_mod_fg) ~ fitted(rich_stab_mod_fg)) # same as usual 
r.squaredGLMM(rich_stab_mod_fg) # 0.819, 0.828
performance::r2(rich_stab_mod_fg)
emtrends(rich_stab_mod_fg, pairwise ~ habitat, var = "functional_richness_mean") 
#  fringing different from all others
# contrast                    estimate     SE  df z.ratio p.value
# Fringing - Backreef           0.1872 0.0468 Inf   3.999  0.0004
# Fringing - Forereef 10m       0.2348 0.0519 Inf   4.524  <.0001
# Fringing - Forereef 17m       0.2040 0.0489 Inf   4.173  0.0002
# Backreef - Forereef 10m       0.0477 0.0472 Inf   1.011  0.7432
# Backreef - Forereef 17m       0.0169 0.0439 Inf   0.385  0.9806
# Forereef 10m - Forereef 17m  -0.0308 0.0475 Inf  -0.648  0.9164
#
```

@@

```{r synchrony staability mod}
synch_stab_mod <- glmmTMB(cover_stability ~ synchrony*habitat + (1|site), family = Gamma("log"), data = diversity_stability_synchrony)
summary(synch_stab_mod)
Anova(synch_stab_mod)
#                     Chisq Df Pr(>Chisq)    
# synchrony         1308.33  1  < 2.2e-16 ***
# habitat            103.13  3  < 2.2e-16 ***
# synchrony:habitat  327.97  3  < 2.2e-16 ***
hist(residuals(synch_stab_mod)) # looks good
plot(residuals(synch_stab_mod) ~ fitted(synch_stab_mod)) # same issue as always
r.squaredGLMM(synch_stab_mod) # 0.63, 0.71
performance::r2(synch_stab_mod)
emtrends(synch_stab_mod, pairwise ~ habitat, var = "synchrony") # fringing and backreef not different
# contrast                    estimate    SE  df z.ratio p.value
# Fringing - Backreef            0.181 0.107 Inf   1.695  0.3262
# Fringing - Forereef 10m        1.492 0.142 Inf  10.487  <.0001
# Fringing - Forereef 17m        2.436 0.149 Inf  16.336  <.0001
# Backreef - Forereef 10m        1.311 0.157 Inf   8.346  <.0001
# Backreef - Forereef 17m        2.255 0.161 Inf  14.008  <.0001
# Forereef 10m - Forereef 17m    0.944 0.181 Inf   5.222  <.0001
```

@@

#                           Chisq Df Pr(>Chisq)   
# avg_richness_fg          3.4019  1   0.065122 . 
# habitat                 12.8194  3   0.005044 **
# avg_richness_fg:habitat  7.9921  3   0.046176 *

hist(residuals(rich_stab_mod_site_fg)) # blocky but fine
plot(residuals(rich_stab_mod_site_fg) ~ fitted(rich_stab_mod_site_fg)) # heteroskedastic
r.squaredGLMM(rich_stab_mod_site_fg) # 0.563
performance::r2(rich_stab_mod_site_fg) # 0.551
emtrends(rich_stab_mod_site_fg, pairwise ~ habitat, var = "functional_richness_mean") 
# Fringing different from outer 10
# contrast                    estimate    SE  df z.ratio p.value
# Fringing - Backreef           0.2552 0.254 Inf   1.005  0.7463
# Fringing - Forereef 10m       1.2611 0.469 Inf   2.686  0.0364
# Fringing - Forereef 17m       0.3543 0.287 Inf   1.235  0.6045
# Backreef - Forereef 10m       1.0059 0.492 Inf   2.044  0.1720
# Backreef - Forereef 17m       0.0991 0.323 Inf   0.307  0.9900
# Forereef 10m - Forereef 17m  -0.9068 0.510 Inf  -1.778  0.2838
```

@@

### model
```{r synchrony staability mod site}
synch_stab_mod_site <- glmmTMB(cover_stability ~ synchrony*habitat, family = Gamma("log"), data = diversity_stability_synchrony_site)
summary(synch_stab_mod_site)
Anova(synch_stab_mod_site)
#                     Chisq Df Pr(>Chisq)    
# synchrony         22.0686  1  2.631e-06 ***
# habitat            9.3333  3    0.02517 *  
# synchrony:habitat 10.5561  3    0.01439 *  
hist(residuals(synch_stab_mod_site)) # almost uniform
plot(residuals(synch_stab_mod_site) ~ fitted(synch_stab_mod_site)) # same issue as always
r.squaredGLMM(synch_stab_mod_site) # 0.678
performance::r2(synch_stab_mod_site) # 0.702
emtrends(synch_stab_mod_site, pairwise ~ habitat, var = "synchrony") # backreef is different from forereef 17, fringing is a p = 0.05
#
# contrast                    estimate    SE  df z.ratio p.value
# Fringing - Backreef           -0.256 0.971 Inf  -0.263  0.9936
# Fringing - Forereef 10m        1.324 0.916 Inf   1.445  0.4709
# Fringing - Forereef 17m        2.930 1.142 Inf   2.567  0.0503
# Backreef - Forereef 10m        1.580 0.858 Inf   1.842  0.2534
# Backreef - Forereef 17m        3.186 1.095 Inf   2.909  0.0190
# Forereef 10m - Forereef 17m    1.606 1.047 Inf   1.534  0.4172
```

@@

# calculate mean synchrony within a site (!= synchrony of species at the site level!)
mean_synchrony_stability <- diversity_stability_synchrony %>% 
  group_by(habitat, site) %>% 
  summarise(synchrony_mean = mean(synchrony, na.rm = T),
            stability_mean = mean(cover_stability, na.rm = T))

@@

(spatial_synchrony_plot <- 
  ggplot() +
  geom_point(data = dss_spatial_2,
             aes(x = spatial_synchrony, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_spat_stab_effects, aes(x = spatial_synchrony, y = exp(yvar), colour = habitat), size = 1.5) +
  geom_ribbon(data = filtered_spat_stab_effects, aes(x = spatial_synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Stability\n", x = "Spatial synchrony", title = "") +
  trends_theme
         )

@@

# ggsave(filename = "output/figure_5_option3.png", figure_5, height = 6, width = 22)
```

# SUPPLEMENTAL FIGURE CODE:

@@

### prep
```{r richness df}
quad_richness_df <- alpha_diversity_quad_macro %>% 
  group_by(year, habitat) %>% 
  summarise(richness_mean = mean(richness),
            richness_se = se(richness))
```

>>>>

cover_means <- cover_df %>%
  dplyr::group_by(year, habitat) %>%
  dplyr::summarise(macroalgae = mean(macroalgae),
            coral = mean(coral),
            turf = mean(turf),
            cca = mean(cca),
            sand = mean(sand),
            other = mean(other)
            )

cover_se <- cover_df %>%
  dplyr::group_by(year, habitat) %>%
  dplyr::summarise(macroalgae = se(macroalgae),
            coral = se(coral),
            turf = se(turf),
            cca = se(cca),
            sand = se(sand),
            other = se(other)
            )
```

@@

taxa_sum_check <- macro_functional_groups %>%
  dplyr::group_by(location, year) %>%
  dplyr::summarise(sum_macrocover = sum(prop_cover))

@@

#summarizing functional groups at the quad level
fg_summary <- macro_functional_groups %>%
  dplyr::group_by(location, year, functional_group, habitat, site) %>%
  dplyr::summarise(sum_fg_cover = sum(prop_cover))

@@

#summarizing functional groups at the site level
#NOTE that this is ADDED and not averaged. This data is only used to find richness at the site level. LE did this as adding and as a sum, both end up the same from a richness perspective. 
fg_summary_site_SUM <- macro_functional_groups %>%
  dplyr::group_by(site, year, habitat, functional_group) %>%
  dplyr::summarise(sum_fg_cover = sum(prop_cover))

#QAQC
fg_summary_qaqc <- macro_functional_groups %>%
  dplyr::group_by(location, year, functional_group) %>%
  dplyr::summarise(sum_fg_cover = sum(prop_cover))

fg_qaqc <- fg_summary_qaqc %>%
  dplyr::group_by(location, year) %>%
  dplyr::summarise(sum_fg_check = sum(sum_fg_cover))

@@

#pivot wide at quad level
fg_summary_wide <- fg_summary %>%
  tidyr::pivot_wider(names_from = "functional_group", values_from = "sum_fg_cover")

@@

cover_mod <- glmmTMB(cover_trans ~ richness*habitat + (1|site/location) + (1|year), family = beta_family(), data = alpha_diversity_quad_macro)
summary(cover_mod)
Anova(cover_mod)
#                     Chisq Df Pr(>Chisq)    
# richness         30918.96  1  < 2.2e-16 ***
# habitat            141.68  3  < 2.2e-16 ***
# richness:habitat   294.44  3  < 2.2e-16 ***
hist(residuals(cover_mod)) # looks fine
plot(residuals(cover_mod) ~ fitted(cover_mod)) # negative trend but it can't be less than 0 so not too concerned. some wave pattern
r.squaredGLMM(cover_mod) # 0.631, 0.650
performance::r2(cover_mod) # 0.635, 0.654
emtrends(cover_mod, pairwise ~ habitat, var = "richness") # backreef not different from forereef, forereef also not different
# contrast                    estimate     SE  df z.ratio p.value
# Fringing - Backreef          0.34715 0.0241 Inf  14.407  <.0001
# Fringing - Forereef 10m      0.34862 0.0240 Inf  14.513  <.0001
# Fringing - Forereef 17m      0.38937 0.0250 Inf  15.593  <.0001
# Backreef - Forereef 10m      0.00147 0.0205 Inf   0.072  0.9999
# Backreef - Forereef 17m      0.04223 0.0214 Inf   1.975  0.1974
# Forereef 10m - Forereef 17m  0.04076 0.0209 Inf   1.947  0.2087

@@

summary(cover_mod_fg)
car::Anova(cover_mod_fg)
#                                 Chisq Df Pr(>Chisq)    
# functional_richness         31362.191  1  < 2.2e-16 ***
# habitat                        49.343  3  1.103e-10 ***
# functional_richness:habitat   451.974  3  < 2.2e-16 ***
hist(residuals(cover_mod_fg)) # good

r.squaredGLMM(cover_mod_fg) # 0.64, 0.65
performance::r2(cover_mod_fg) # 0.63, 0.65
emtrends(cover_mod_fg, pairwise ~ habitat, var = "functional_richness") 
# contrast                    estimate     SE  df z.ratio p.value
# Fringing - Backreef            0.211 0.0266 Inf   7.937  <.0001
# Fringing - Forereef 10m        0.348 0.0262 Inf  13.285  <.0001
# Fringing - Forereef 17m        0.531 0.0262 Inf  20.280  <.0001
# Backreef - Forereef 10m        0.136 0.0231 Inf   5.888  <.0001
# Backreef - Forereef 17m        0.320 0.0232 Inf  13.812  <.0001
# Forereef 10m - Forereef 17m    0.184 0.0223 Inf   8.253  <.0001

@@

```{r synchrony cover mod}
synch_cov_mod <- glmmTMB(cover_mean ~ synchrony*habitat + (1|site), family = beta_family(), data = diversity_stability_synchrony)
summary(synch_cov_mod)
Anova(synch_cov_mod)
#                     Chisq Df Pr(>Chisq)    
# synchrony         209.521  1  < 2.2e-16 ***
# habitat            52.498  3  2.345e-11 ***
# synchrony:habitat  54.627  3  8.246e-12 ***
hist(residuals(synch_cov_mod)) # slight skew
plot(residuals(synch_cov_mod) ~ fitted(synch_cov_mod)) # same issues as always

# can't compute r2 - manually fit null model. Null model won't fit without sv_trans() for some reason
null_synch_mod <- glmmTMB(sv_trans(cover_mean) ~ 1, family = beta_family(), data = diversity_stability_synchrony)
performance::r2(model = synch_cov_mod, null_model = null_synch_mod) # 0.352, 0.534
emtrends(synch_cov_mod, pairwise ~ habitat, var = "synchrony") # forereef 17 distinct from all other habitats

@@

### model at site level
```{r site richness mod}
site_cover_mod <- glmmTMB(cover_trans ~ richness*habitat + (1|site_habitat) + (1|year), family = beta_family(), data = alpha_diversity_site_macro)
summary(site_cover_mod)
Anova(site_cover_mod)
#                    Chisq Df Pr(>Chisq)    
# richness         58.9196  1  1.642e-14 ***
# habitat           6.6248  3    0.08487 .  
# richness:habitat  3.2966  3    0.34811    
hist(residuals(site_cover_mod)) # some positive skew, not the worst thing ever
plot(residuals(site_cover_mod) ~ fitted(site_cover_mod)) # a little heteroscedastic but can't be less than 0
r.squaredGLMM(site_cover_mod) # 0.2212534, 0.5990825
performance::r2(site_cover_mod) # 0.22, 0.59
emtrends(site_cover_mod, pairwise ~ habitat, var = "richness") # no difference between the habitats
# contrast                    estimate     SE  df z.ratio p.value
# Fringing - Backreef          -0.0163 0.0394 Inf  -0.413  0.9762
# Fringing - Forereef 10m       0.0639 0.0503 Inf   1.271  0.5818
# Fringing - Forereef 17m       0.0334 0.0612 Inf   0.546  0.9476
# Backreef - Forereef 10m       0.0801 0.0460 Inf   1.742  0.3020
# Backreef - Forereef 17m       0.0497 0.0583 Inf   0.853  0.8289
# Forereef 10m - Forereef 17m  -0.0304 0.0651 Inf  -0.467  0.9662
```

@@

#                               Chisq Df Pr(>Chisq)    
# functional_richness         30.3776  1  3.556e-08 ***
# habitat                      4.9314  3     0.1769    
# functional_richness:habitat  4.1429  3     0.2464

hist(residuals(site_cover_mod_fg)) # positive skew
plot(residuals(site_cover_mod_fg) ~ fitted(site_cover_mod_fg)) # a little heteroscedastic but can't be less than 0
r.squaredGLMM(site_cover_mod_fg) # 0.180, 0.563
performance::r2(site_cover_mod_fg) # 0.180, 0.563
emtrends(site_cover_mod_fg, pairwise ~ habitat, var = "functional_richness") # no difference between the habitats
# contrast                    estimate     SE  df z.ratio p.value
# Fringing - Backreef          -0.0163 0.0394 Inf  -0.413  0.9762
# Fringing - Forereef 10m       0.0639 0.0503 Inf   1.271  0.5818
# Fringing - Forereef 17m       0.0334 0.0612 Inf   0.546  0.9476
# Backreef - Forereef 10m       0.0801 0.0460 Inf   1.742  0.3020
# Backreef - Forereef 17m       0.0497 0.0583 Inf   0.853  0.8289
# Forereef 10m - Forereef 17m  -0.0304 0.0651 Inf  -0.467  0.9662

@@

```{r synchrony cover mod}
synch_cov_mod_site <- glmmTMB(cover_mean ~ synchrony*habitat, family = beta_family(), data = diversity_stability_synchrony_site)
summary(synch_cov_mod_site)
Anova(synch_cov_mod_site)
#                    Chisq Df Pr(>Chisq)  
# synchrony         2.4156  1    0.12013  
# habitat           7.7966  3    0.05041 .
# synchrony:habitat 8.8876  3    0.03082 *
performance::r2(synch_cov_mod_site) # 0.48
hist(residuals(synch_cov_mod_site)) # slight skew
plot(residuals(synch_cov_mod_site) ~ fitted(synch_cov_mod_site)) # looks fine
emtrends(synch_cov_mod_site, pairwise ~ habitat, var = "synchrony") # forereef 17 different from backreef
# contrast                    estimate   SE  df z.ratio p.value
# Fringing - Backreef            -4.31 3.39 Inf  -1.269  0.5824
# Fringing - Forereef 10m        -0.29 3.23 Inf  -0.090  0.9997
# Fringing - Forereef 17m         3.66 3.79 Inf   0.968  0.7679
# Backreef - Forereef 10m         4.02 1.98 Inf   2.031  0.1764
# Backreef - Forereef 17m         7.97 2.80 Inf   2.848  0.0228
# Forereef 10m - Forereef 17m     3.95 2.59 Inf   1.526  0.4218
```

@@

```{r richness stability mod}
rich_stab_mod <- glmmTMB(cover_stability~ richness_mean*habitat + (1|site), family = Gamma(link = "log"), data = diversity_stability_synchrony)
summary(rich_stab_mod)
Anova(rich_stab_mod)
#                          Chisq Df Pr(>Chisq)    
# richness_mean         2720.311  1  < 2.2e-16 ***
# habitat                325.657  3  < 2.2e-16 ***
# richness_mean:habitat   33.857  3  2.124e-07 ***
hist(residuals(rich_stab_mod)) # looks good
plot(residuals(rich_stab_mod) ~ fitted(rich_stab_mod)) # heteroskedastic - bring to group
r.squaredGLMM(rich_stab_mod) # 0.803, 0.808
performance::r2(rich_stab_mod) # 0.812, 0.816
emtrends(rich_stab_mod, pairwise ~ habitat, var = "richness_mean") # fringing not diff from outer17, backreef not different from outer10
# $contrasts
#  contrast                     estimate     SE  df z.ratio p.value
#  Fringing - Backreef          0.186640 0.0414 Inf   4.511  <.0001
#  Fringing - Forereef 10m      0.187248 0.0455 Inf   4.114  0.0002
#  Fringing - Forereef 17m      0.027482 0.0463 Inf   0.594  0.9340
#  Backreef - Forereef 10m      0.000608 0.0405 Inf   0.015  1.0000
#  Backreef - Forereef 17m     -0.159158 0.0411 Inf  -3.875  0.0006
#  Forereef 10m - Forereef 17m -0.159766 0.0439 Inf  -3.640  0.0016

@@

#                                    Chisq Df Pr(>Chisq)    
# functional_richness_mean         3082.22  1  < 2.2e-16 ***
# habitat                           116.63  3  < 2.2e-16 ***
# functional_richness_mean:habitat   25.79  3  1.055e-05 ***

hist(residuals(rich_stab_mod_fg)) # looks fine
plot(residuals(rich_stab_mod_fg) ~ fitted(rich_stab_mod_fg)) # same as usual 
r.squaredGLMM(rich_stab_mod_fg) # 0.819, 0.828
performance::r2(rich_stab_mod_fg)
emtrends(rich_stab_mod_fg, pairwise ~ habitat, var = "functional_richness_mean") 
#  fringing different from all others
# contrast                    estimate     SE  df z.ratio p.value
# Fringing - Backreef           0.1825 0.0468 Inf   3.903  0.0006
# Fringing - Forereef 10m       0.2372 0.0518 Inf   4.575  <.0001
# Fringing - Forereef 17m       0.2045 0.0488 Inf   4.187  0.0002
# Backreef - Forereef 10m       0.0547 0.0472 Inf   1.158  0.6534
# Backreef - Forereef 17m       0.0219 0.0439 Inf   0.500  0.9592
# Forereef 10m - Forereef 17m  -0.0327 0.0475 Inf  -0.689  0.9015
#
```

@@

```{r synchrony staability mod}
synch_stab_mod <- glmmTMB(cover_stability ~ synchrony*habitat + (1|site), family = Gamma("log"), data = diversity_stability_synchrony)
summary(synch_stab_mod)
Anova(synch_stab_mod)
#                     Chisq Df Pr(>Chisq)    
# synchrony         1308.34  1  < 2.2e-16 ***
# habitat            103.13  3  < 2.2e-16 ***
# synchrony:habitat  327.96  3  < 2.2e-16 ***
hist(residuals(synch_stab_mod)) # looks good
plot(residuals(synch_stab_mod) ~ fitted(synch_stab_mod)) # same issue as always
r.squaredGLMM(synch_stab_mod) # 0.63, 0.71
performance::r2(synch_stab_mod) # 0.66, 0.74
emtrends(synch_stab_mod, pairwise ~ habitat, var = "synchrony") # fringing and backreef not different
# contrast                    estimate    SE  df z.ratio p.value
# Fringing - Backreef            0.181 0.107 Inf   1.695  0.3262
# Fringing - Forereef 10m        1.492 0.142 Inf  10.487  <.0001
# Fringing - Forereef 17m        2.436 0.149 Inf  16.336  <.0001
# Backreef - Forereef 10m        1.311 0.157 Inf   8.346  <.0001
# Backreef - Forereef 17m        2.255 0.161 Inf  14.008  <.0001
# Forereef 10m - Forereef 17m    0.944 0.181 Inf   5.222  <.0001
```

@@

# 
#                                    Chisq Df Pr(>Chisq)   
# functional_richness_mean          3.3605  1    0.06678 . 
# habitat                          12.9894  3    0.00466 **
# functional_richness_mean:habitat  7.9219  3    0.04765 *

hist(residuals(rich_stab_mod_site_fg)) # blocky but fine
plot(residuals(rich_stab_mod_site_fg) ~ fitted(rich_stab_mod_site_fg)) # heteroskedastic
r.squaredGLMM(rich_stab_mod_site_fg) # 0.563
performance::r2(rich_stab_mod_site_fg) # 0.551
emtrends(rich_stab_mod_site_fg, pairwise ~ habitat, var = "functional_richness_mean") 
# Fringing different from outer 10
#$contrasts
# contrast                    estimate    SE  df z.ratio p.value
# Fringing - Backreef            0.244 0.252 Inf   0.966  0.7686
# Fringing - Forereef 10m        1.250 0.469 Inf   2.663  0.0388
# Fringing - Forereef 17m        0.354 0.280 Inf   1.264  0.5860
# Backreef - Forereef 10m        1.006 0.493 Inf   2.040  0.1735
# Backreef - Forereef 17m        0.110 0.318 Inf   0.346  0.9858
# Forereef 10m - Forereef 17m   -0.896 0.508 Inf  -1.763  0.2911
```

@@

### model
```{r synchrony staability mod site}
synch_stab_mod_site <- glmmTMB(cover_stability ~ synchrony*habitat, family = Gamma("log"), data = diversity_stability_synchrony_site)
summary(synch_stab_mod_site)
Anova(synch_stab_mod_site)
#                    Chisq Df Pr(>Chisq)    
# synchrony         22.117  1  2.565e-06 ***
# habitat            9.345  3    0.02504 *  
# synchrony:habitat 10.534  3    0.01453 *   
hist(residuals(synch_stab_mod_site)) # almost uniform
plot(residuals(synch_stab_mod_site) ~ fitted(synch_stab_mod_site)) # same issue as always
r.squaredGLMM(synch_stab_mod_site) # 0.678
performance::r2(synch_stab_mod_site) # 0.702
emtrends(synch_stab_mod_site, pairwise ~ habitat, var = "synchrony") # backreef is different from forereef 17, fringing is a p = 0.05
#
# contrast                    estimate    SE  df z.ratio p.value
# Fringing - Backreef           -0.262 0.972 Inf  -0.269  0.9932
# Fringing - Forereef 10m        1.318 0.918 Inf   1.437  0.4764
# Fringing - Forereef 17m        2.924 1.143 Inf   2.559  0.0513
# Backreef - Forereef 10m        1.580 0.857 Inf   1.843  0.2533
# Backreef - Forereef 17m        3.186 1.095 Inf   2.910  0.0189
# Forereef 10m - Forereef 17m    1.606 1.047 Inf   1.534  0.4170
```

@@

# calculate mean synchrony within a site (!= synchrony of species at the site level!)
mean_synchrony_stability <- diversity_stability_synchrony %>% 
  dplyr::group_by(habitat, site) %>% 
  dplyr::summarise(synchrony_mean = mean(synchrony, na.rm = T),
            stability_mean = mean(cover_stability, na.rm = T))

@@

(spatial_synchrony_plot <- 
  ggplot() +
  geom_point(data = dss_spatial_2,
             aes(x = spatial_synchrony, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_spat_stab_effects, aes(x = spatial_synchrony, y = exp(yvar), colour = habitat), size = 1.5) +
  geom_ribbon(data = filtered_spat_stab_effects, aes(x = spatial_synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "", x = "Spatial synchrony", title = "") +
  trends_theme
         )

@@

# ggsave(filename = "output/figure_5_option3_v2.png", figure_5, height = 6, width = 22)
```

@@

### prep
```{r richness df}
quad_richness_df <- alpha_diversity_quad_macro %>% 
  dplyr::group_by(year, habitat) %>% 
  dplyr::summarise(richness_mean = mean(richness),
            richness_se = se(richness))
```
